\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

Large-scale connectomics (the mapping of neural circuits from volumetric
electron microscopy (EM) and X-ray nano-holotomography (XNH) data) requires
automated segmentation of three-dimensional image volumes into individual neuronal
fragments. State-of-the-art automated segmenters produce fragments with high accuracy
but inevitably introduce \emph{split errors}, where a single neuron is represented as
multiple disconnected fragments. Correcting these errors at the scale of modern
connectomics datasets (tens to hundreds of teravoxels) demands automated
proofreading pipelines that can detect and propose corrections for split errors
with high recall and with explicit handling of uncertainty.

This report presents a modular, conservative post-segmentation pipeline that
takes the output of an automated segmenter and produces a graph-based
connectivity representation with confidence-annotated merge proposals. The pipeline
operates in six stages: fragment extraction with TEASAR skeletonization, graph
construction via a skeleton-node KD-tree index, candidate generation with a
four-factor composite scoring model, conservative three-outcome rule-based
validation (ACCEPT / REJECT / AMBIGUOUS), structure assembly, and multi-format
export. Critically, the pipeline treats uncertainty as a first-class output:
ambiguous connections are preserved for human review rather than forced to a
binary decision.

Two key innovations drive the approach's effectiveness on axon-dominated datasets.
First, a \emph{skeleton-node graph} indexes every skeleton node from every fragment
(rather than only degree-1 endpoints), exposing interior splits along long axons
that an endpoint-only graph cannot detect. Second, \emph{distance-conditioned
scoring} switches from a proximity-weighted composite to a proximity-free,
alignment-continuity-weighted score for long-range candidate pairs, preventing
the artificial suppression of genuine long-range splits.

Evaluated on the CREMI Drosophila EM benchmark, the pipeline achieves
Precision = 1.000, Recall = 0.909, and F1 = 0.952. On the XPRESS mouse white
matter XNH challenge, the primary domain-appropriate benchmark, the best
configuration (Experiment~24: direction-weighted scoring, MinGapRule disabled,
SizeDiscrepancyRule relaxed) achieves Coverage = 78.7\% and Recall = 0.9958 on
the training volume, with only 5 false negatives all attributable to the
CurvatureRule. Applied unchanged to the fully held-out validation volume,
the pipeline achieves Coverage = 81.3\% and Recall = 0.9879 (2 false negatives),
confirming that the tuned thresholds generalize without overfitting. The pipeline
is validated by 428 automated tests and includes a reproducible visual diagnostic
report for per-candidate inspection.

The main open challenge is precision: approximately 349{,}000 false-positive
merge proposals are generated alongside 1{,}175 true positives at XPRESS scale
(Precision $\approx$ 0.003), reflecting the difficulty of discriminating
same-axon from different-axon long-range pairs in myelinated white matter using
geometric features alone. A machine-learned post-filter achieves a $10\times$
precision gain at recall 0.85, but cannot maintain recall $\geq 0.99$ with
meaningful precision improvement.
