This work presents a conservative, modular post-segmentation pipeline for
automated split-error detection in connectomics data. The central research
question was: \emph{can a geometry-only, rule-based pipeline achieve high recall
on neural split detection without introducing excessive merge errors?}

On the CREMI Drosophila EM benchmark, the answer is clearly affirmative:
Precision = 1.000, Recall = 0.909, F1 = 0.952, with zero false positives. On
the XPRESS mouse white matter XNH challenge, a harder domain where the
segmented structures are elongated tubular axons with interior splits, the
best configuration (Experiment~24) achieves Recall = 0.9958 on the training
oracle and Recall = 0.9879 on the fully held-out validation volume, confirming
generalization.

\paragraph{Key findings.}
Two innovations are responsible for the majority of the XPRESS recall improvement:

\begin{enumerate}
  \item The \textbf{skeleton-node graph} raises oracle coverage by 56.8
        percentage points over an endpoint-only graph in a single change, by
        enabling detection of interior splits along long axons.

  \item \textbf{Distance-conditioned scoring} prevents the proximity decay from
        suppressing long-range candidates whose alignment and continuity are
        strong, enabling the pipeline to accept splits at gaps up to 2000~nm.
\end{enumerate}

\paragraph{Challenges.}
The dominant open challenge is precision: Precision $\approx 0.003$ on XPRESS,
meaning approximately 297 false-positive merge proposals for every true positive.
This stems from the geometric feature space being insufficient to discriminate
same-axon from cross-axon long-range pairs at the density of myelinated white
matter. A machine-learned filter (Experiment~20) improved precision $10\times$
(to 0.037) at the cost of recall, but could not maintain recall $\geq 0.99$.

The MinGapRule challenge was resolved in Experiment~24: disabling the rule
(min\_gap\_nm = 0) recovered all 37 short-gap false negatives and raised
training recall from 0.966 to 0.9958. The only remaining false negatives
(5 on training, 2 on validation) are caused by the CurvatureRule at gaps of
66--987~nm, where the local direction estimate is unreliable.

The coverage ceiling imposed by the 2000~nm search radius means at least 247
oracle pairs (with gap or PCA-tip distance exceeding this threshold) cannot be
reached by the current graph construction. The best achieved coverage is 78.7\%
on the training volume.

\paragraph{Summary of contributions.}
The pipeline is fully implemented, tested (428 tests, 0 failures), documented,
and open-source. It includes a YAML configuration system, six processing stages,
multi-format export, a ground-truth evaluation framework for two benchmarks,
an ML-based post-filter, and a visual diagnostic report generator. Twenty-five
experiments with detailed quantitative logs trace the development of each component.
