\section{Completed: Experiment 24 (Direction Weights and MinGapRule Removal)}

Experiment~24 has been completed. Disabling the MinGapRule (\texttt{min\_gap\_nm} = 0)
recovered all 37 short-gap false negatives and 114 previously uncovered oracle
pairs. Simultaneously, raising \texttt{max\_radius\_ratio} from 15.0 to 20.0
recovered 2 SizeDiscrepancyRule false negatives, and adjusting short-range
scoring weights (alignment 0.30$\to$0.45, proximity 0.35$\to$0.15) improved
discrimination of well-aligned pairs. The final training result is
Recall = 0.9958 (FN = 5, all CurvatureRule) and the held-out validation result
is Recall = 0.9879 (FN = 2, both CurvatureRule). Experiment~25, which tested
raising \texttt{skip\_distance\_nm} from 1000 to 1100~nm in the CurvatureRule,
was negative (Recall 0.9958$\to$0.9949) and was reverted.

The 5 remaining false negatives all fail the CurvatureRule at gaps of 66--987~nm
and have composite scores of 0.41--0.64. Because the direction estimate at short
range is sensitive to skeleton geometry near the gap boundary, these cases appear
unrecoverable without access to raw image data to disambiguate curvature.

\section{Precision Improvement}

Addressing the precision gap (currently $\approx$0.003) is the highest-priority research
problem. Three approaches are under consideration:

\begin{enumerate}
  \item \textbf{Improved geometric features.} The current features do not
        capture local neighborhood density (how many other fragments are nearby
        a proposed pair) or myelin sheath consistency. Adding a neighborhood
        density feature may discriminate same-axon pairs (isolated in their local
        neighborhood) from cross-axon pairs (surrounded by many other axons).

  \item \textbf{Retrained ML filter.} The Experiment~20 classifier was trained
        on degree features that are effective at recall 0.85 but not at 0.99.
        A model trained specifically to maximize precision at recall $\geq 0.99$
        (using a custom loss function or recall-constrained Pareto search) may
        achieve better results.

  \item \textbf{Agglomerative confidence scoring.} Rather than accepting all
        candidates above a fixed composite threshold, a global agglomerative
        approach could select the best non-conflicting set of merges across the
        entire volume, trading some recall for large precision gains.
\end{enumerate}

\section{Coverage Expansion}

The current 247 unreachable oracle pairs fall into two categories: 172 with
true gap $> 2000$~nm, and 75 involving PCA-only endpoint fragments whose
bounding-box tip-to-tip distance does not span the gap. Two options:

\begin{itemize}
  \item Raise \texttt{max\_endpoint\_search\_nm} from 2000 to 3000~nm.
        Experiment~13 showed this causes recall regression, but that was before
        distance-conditioned scoring (Exp~14) and CurvatureRule skip (Exp~16)
        were added. A re-test at 3000~nm with the current scoring may give
        different results.

  \item Raise \texttt{max\_skeleton\_voxels} to cover the 63 PCA-only fragments,
        each requiring 5--30 seconds of TEASAR time ($\sim 13$ minutes additional
        runtime total). Better skeleton coverage would improve endpoint accuracy
        for large fragments.
\end{itemize}

\section{XPRESS Challenge Submission}

The pipeline results have not yet been submitted to the XPRESS challenge portal
(\url{https://xpress.grand-challenge.org/}) for official evaluation. A formal
submission would provide an external, blind evaluation and place the results
in the context of other participating systems.

\section{Final Presentation}

This document is the final written report. The remaining deliverable is the
end-of-semester presentation, which will cover the pipeline design, the
progression of experiments, and the final quantitative results on both the
CREMI and XPRESS benchmarks. Key items to highlight include the skeleton-node
graph construction, distance-conditioned scoring, and the generalization results
from the held-out validation volume.
