\section{Design Philosophy}

The pipeline is built around three principles. First, \textbf{conservative
correctness}: a merge that should not happen is worse than a merge proposal
that is correctly rejected, so the pipeline never forces an ambiguous decision;
uncertainty is output explicitly as a third outcome (AMBIGUOUS). Second,
\textbf{modularity}: each stage can be run independently, with intermediate
results saved to disk and re-used across parameter sweeps. Third,
\textbf{reproducibility}: all parameters are controlled through a single YAML
configuration file, and random seeds are fixed, so the same config always
produces the same output.

The six-stage pipeline is shown in Figure~\ref{fig:pipeline}. The remainder
of this chapter describes each stage.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  node distance = 4.5mm,
  every node/.style  = {align=center},
  %--- node styles ---
  io/.style = {
    draw, ellipse,
    minimum width = 62mm, minimum height = 9mm,
    fill = gray!12, draw = gray!55, line width = 0.65pt,
    font = \small\bfseries,
  },
  stage/.style = {
    draw, rounded corners = 5pt,
    minimum width = 90mm,
    inner xsep = 10pt, inner ysep = 7pt,
    fill = blue!7, draw = blue!35!black, line width = 0.65pt,
  },
  key/.style = {stage, fill = orange!7, draw = orange!40!black},
  out/.style = {stage, fill = green!7, draw = green!45!black},
  %--- arrow style ---
  arr/.style = {
    -{Stealth[length=5pt, width=4pt]},
    line width = 0.9pt, color = gray!60!black,
  },
]

%% Input node
\node[io] (inp) {Segmentation Volume};

%% Stage 1
\node[stage, below=of inp] (s1) {
  {\small\bfseries [1]\enspace Fragment Extraction}\\[2pt]
  {\footnotesize\color{gray!55!black}
    Connected-component extraction $\cdot$ TEASAR skeletonisation $\cdot$ chunk stitching}
};

%% Stage 2
\node[stage, below=of s1] (s2) {
  {\small\bfseries [2]\enspace Graph Construction}\\[2pt]
  {\footnotesize\color{gray!55!black}
    Skeleton-node KD-tree $\cdot$ optional long-range endpoint pass}
};

%% Stage 3
\node[stage, below=of s2] (s3) {
  {\small\bfseries [3]\enspace Candidate Generation}\\[2pt]
  {\footnotesize\color{gray!55!black}
    Composite scoring $\cdot$ distance-conditioned weights}
};

%% Stage 4 — highlighted as the key decision stage
\node[key, below=of s3] (s4) {
  {\small\bfseries [4]\enspace Conservative Validation}\\[2pt]
  {\footnotesize\color{gray!55!black}
    7 configurable rules $\cdot$ three outcomes: \textbf{accept} / \textbf{reject} / \textbf{ambiguous}}
};

%% Stage 5
\node[stage, below=of s4] (s5) {
  {\small\bfseries [5]\enspace Assembly}\\[2pt]
  {\footnotesize\color{gray!55!black}
    Connected-component merging $\cdot$ cycle detection $\cdot$ topology checks}
};

%% Stage 6 — output
\node[out, below=of s5] (s6) {
  {\small\bfseries [6]\enspace Export}\\[2pt]
  {\footnotesize\color{gray!55!black}
    CSV $\cdot$ GraphML $\cdot$ SWC $\cdot$ Neuroglancer annotations}
};

%% Arrows
\draw[arr] (inp) -- (s1);
\draw[arr] (s1)  -- (s2);
\draw[arr] (s2)  -- (s3);
\draw[arr] (s3)  -- (s4);
\draw[arr] (s4)  -- (s5);
\draw[arr] (s5)  -- (s6);

\end{tikzpicture}
\caption{Overview of the six-stage post-segmentation pipeline. Stage~4
(Conservative Validation, shaded orange) is the primary decision point; it
returns one of three explicit outcomes so that uncertainty is never silently
discarded.}
\label{fig:pipeline}
\end{figure}

\section{Stage 1: Fragment Extraction}

\paragraph{Connected component extraction.}
The input is a voxel-wise label volume (HDF5 format). Because full volumes
may exceed available memory ($699^3$ voxels at 64 bits $\approx$ 2.7~GB), the
pipeline reads the volume in configurable chunks with overlap, extracts
connected components per chunk, and stitches cross-boundary fragments using
label agreement in overlap zones.

Each resulting \texttt{Fragment} object stores its integer label identifier,
voxel count, axis-aligned bounding box, centroid (in physical nm coordinates),
a list of endpoint coordinates, an optional TEASAR skeleton, and a boundary flag
indicating whether the fragment touches a chunk boundary.

\paragraph{Skeletonization.}
Fragments above a minimum voxel threshold (100 voxels by default, configurable)
are skeletonized using the \texttt{kimimaro} TEASAR implementation. The TEASAR
invalidation radius parameter \texttt{invalidation\_d0} controls skeleton
branch density; we use \texttt{invalidation\_d0 = 10} for XPRESS.

For very large fragments (above a configurable \texttt{max\_skeleton\_voxels}
threshold, set to 500{,}000 voxels for XPRESS), TEASAR runtime scales
non-linearly and can require $>10$ minutes per fragment. For these, the pipeline
falls back to a PCA-based bounding-box endpoint: the fragment's bounding box
is projected onto its first principal component, and the two extreme corners
of the projected box are used as pseudo-endpoints.

\paragraph{Fragment stitching.}
The stitching module uses a union-find data structure to merge fragment IDs
across chunk boundaries, relabeling fragments globally after processing all
chunks. On the XPRESS training volume, 16{,}187 per-chunk fragments stitch
into 12{,}935 global fragments.

\section{Stage 2: Graph Construction}

Graph construction determines which fragment pairs are considered as merge
candidates. This stage had the single largest impact on recall in our
experiments.

\paragraph{Endpoint graph (baseline).}
The naive approach indexes only TEASAR degree-1 endpoints (skeleton leaf nodes)
in a KD-tree and connects pairs within a distance threshold
(\texttt{max\_distance\_nm}). This correctly handles splits at the tips of axons
but cannot represent splits in the interior of long axons, where the split point
is surrounded by skeleton interior nodes, not endpoints.

\paragraph{Skeleton-node graph (proposed method).}
The \texttt{skeleton\_node} construction method indexes \emph{every} skeleton
node (leaf, branch, and interior) from every fragment in a single global
KD-tree. All node-to-node pairs within \texttt{max\_distance\_nm} receive an
edge. This exposes interior splits at any point along an axon.

On XPRESS, switching from the endpoint graph (at 300-voxel crop resolution) to
the skeleton-node graph (at full-volume 699-voxel resolution) raised oracle
coverage from 23\% to 79.8\%, a 56.8-percentage-point gain in one change.

\paragraph{Long-range endpoint pass.}
A supplemental second pass queries only degree-1 endpoints at a larger radius
(\texttt{max\_endpoint\_search\_nm} = 2000~nm) using the same KD-tree.
This adds edges for genuine segmentation gaps that are wider than the skeleton-node
radius (500~nm), targeting true splits where the two fragments' interior skeleton
nodes are not close but their free endpoints point toward each other across the gap.

\section{Stage 3: Candidate Generation and Scoring}

\paragraph{Composite score.}
For each graph edge $(A, B)$, the candidate generator identifies the closest
endpoint pair using the skeleton node positions, then computes four normalized
sub-scores:

\begin{description}
  \item[Proximity score] $s_\text{prox} = \exp(-d / d_\text{ref})$, where $d$
        is the gap distance and $d_\text{ref}$ is a configurable reference
        distance (600~nm for XPRESS). Decays to zero as distance grows.

  \item[Alignment score] $s_\text{align} = (\hat{t}_A \cdot \hat{t}_B + 1) / 2$,
        where $\hat{t}_A$ and $\hat{t}_B$ are the unit tangent vectors at the
        endpoints of fragments $A$ and $B$, estimated from the local skeleton
        geometry. A score of 1.0 means the fragments point directly toward each
        other; 0.0 means they are antiparallel.

  \item[Continuity score] $s_\text{cont} = \max(0,\; 1 - \kappa / \kappa_\text{max})$,
        where $\kappa$ is the angular bend at the proposed junction and
        $\kappa_\text{max} = 180°$. Penalizes sharp turns.

  \item[Size score] $s_\text{size} = \min(r_A, r_B) / \max(r_A, r_B)$, where
        $r_A, r_B$ are the estimated radii (cube-root of voxel count). Penalizes
        large radius discrepancies.
\end{description}

The composite score is a weighted sum:
\begin{equation}
  s_\text{comp} = w_\text{prox} s_\text{prox} + w_\text{align} s_\text{align}
                + w_\text{cont} s_\text{cont} + w_\text{size} s_\text{size}
\end{equation}
with default weights $(0.35, 0.30, 0.25, 0.10)$ for the four components.

\paragraph{Distance-conditioned scoring.}
For long-range pairs (gap $> \texttt{long\_range\_threshold\_nm}$ = 1000~nm),
the proximity score decays below 0.05, which dominates the composite and
suppresses the score even when alignment and continuity are strong. To address
this, the pipeline switches to a proximity-free weight vector
$(0.00, 0.45, 0.40, 0.15)$ for long-range pairs, emphasizing alignment and
continuity. This was introduced in Experiment~14 and raised recall from
0.898 to 0.889 on the short-range subset and from a much lower baseline on
the long-range subset.

\section{Stage 4: Conservative Validation}

\paragraph{Three-outcome model.}
Every candidate is assigned one of three outcomes: ACCEPT (high confidence the
merge is correct), REJECT (high confidence it is not), or AMBIGUOUS (insufficient
evidence). Ambiguous candidates are preserved in the output and excluded from
precision/recall computation; they represent explicit uncertainty, not errors.

\paragraph{Validation rules.}
Seven rules are applied in sequence; any REJECT terminates evaluation of
that candidate:

\begin{enumerate}
  \item \textbf{MinGapRule}: Rejects candidates with gap $< \texttt{min\_gap\_nm}$.
        Intended to filter adjacency false positives (different axons sharing a
        voxel boundary) but also rejects genuine short-gap splits (Experiment~22
        shows this rule causes 37 false negatives in the current config).

  \item \textbf{MaxDistanceRule}: Hard-rejects if the endpoint-to-endpoint gap
        exceeds \texttt{max\_distance\_nm} (set to 20{,}000~nm for XPRESS to
        accommodate PCA-based endpoint pairs, whose tip-to-tip distance can reach
        12{,}000~nm).

  \item \textbf{CurvatureRule}: Rejects if the junction angle exceeds
        \texttt{max\_curvature\_deg} (150°). For long-range pairs
        (gap $> \texttt{skip\_distance\_nm}$ = 1000~nm), the rule is skipped
        entirely, because the endpoint-centroid direction estimate is unreliable
        at large distances and causes false rejections.

  \item \textbf{DirectionReversalRule}: Rejects if the two fragment tangent
        vectors point away from each other (dot product $< 0$).

  \item \textbf{SizeDiscrepancyRule}: Rejects if the cube-root radius ratio
        exceeds \texttt{max\_radius\_ratio} (15.0 for XPRESS).

  \item \textbf{BranchingLimitRule}: Rejects if the merge would connect to more
        than \texttt{max\_branches} (5{,}000) already-accepted partners. This is
        a safeguard against pathological graph nodes.

  \item \textbf{CompositeScoreRule}: Hard-rejects if the composite score is below
        \texttt{reject\_threshold} (0.15).
\end{enumerate}

All rules are configurable via YAML and may be omitted or reordered.

\paragraph{Accept/reject thresholds.}
After all rules pass, a candidate is accepted if its composite score exceeds
\texttt{accept\_threshold} (0.25). Candidates between \texttt{reject\_threshold}
and \texttt{accept\_threshold} are marked AMBIGUOUS.

\section{Stage 5: Assembly}

Accepted connections are added as edges to the fragment graph. Connected
components of this graph define assembled structures (reconstructed neurons
or axon segments). The assembler checks each structure for cycles, excessive
branching, and ambiguous connections, attaching topology warnings as metadata.
Per-structure confidence is computed as the minimum composite score across all
accepted connections (weakest-link model).

\section{Stage 6: Export}

The pipeline exports results in up to five formats: CSV tables of fragment
metadata and per-connection decisions (default, used for all XPRESS experiments),
GraphML (disabled by default due to OOM risk at scale, as 815{,}000+ edges produce
multi-gigabyte XML files), SWC neuron morphology, Neuroglancer annotation layers,
and a corrected precomputed segmentation volume with accepted merges applied.

\section{Ground Truth Evaluation}

\paragraph{CREMI oracle.}
For CREMI, the ground truth is available as human-annotated segment labels. The
evaluation oracle considers two fragments as a true merge pair if they share the
same ground-truth label ID. Evaluation computes precision, recall, and F1 over
accepted candidates, excluding ambiguous candidates from all denominators.

\paragraph{XPRESS skeleton oracle.}
For XPRESS, ground truth is provided as manually traced skeleton graphs
(\texttt{XPRESS\_training\_skels.npz}). For each ground-truth axon trajectory,
the two segmentation fragments containing the most skeleton nodes on either side
of a trajectory gap define a merge pair. The oracle is built by:
(1) mapping skeleton node coordinates (in physical nm space) to segmentation
label IDs by voxel lookup, accounting for the volume's physical offset;
(2) identifying the dominant label for each trajectory segment; and
(3) recording pairs of labels that are separated by a gap in the ground-truth
trajectory but belong to the same axon.

Two metrics are reported: \emph{Coverage} (fraction of oracle pairs for which
the pipeline generates at least one candidate), measuring graph-construction
reach, and \emph{Recall} (fraction of covered pairs that are accepted),
measuring validation accuracy. Precision is computed as the fraction of accepted
candidates that correspond to oracle pairs.

\section{Visual Validation Report}
\label{sec:visual-validation}

To support qualitative inspection of pipeline behavior, we developed a diagnostic
report generator (\texttt{scripts/generate\_visual\_report.py}) that produces a
multi-page PDF or inline GitHub images.  For a stratified sample of candidates
(high-confidence accepted, low-confidence accepted, borderline rejected, oracle
false negatives, and random rejected), each candidate occupies one row of
three panels side-by-side:

\begin{enumerate}
  \item \textbf{Raw / Neutral:} Grayscale EM image (when available) or a
        neutral segmentation view in which each segment receives a distinct
        muted color but neither fragment is highlighted. Ground-truth skeleton
        nodes and edges near the Z-slice are overlaid as yellow dots/lines,
        showing where the manually traced neuron paths run through the tissue.
        This panel provides tissue context without the pipeline's coloring
        choices.

  \item \textbf{Pipeline Prediction:} Colorized segmentation with Fragment~A
        in red and Fragment~B in blue. A white arrow indicates the proposed
        connection direction; an ACC/REJ badge (green/red) states the decision.
        All five component scores (proximity, alignment, continuity, size,
        composite) and the gap distance are shown in the panel title.

  \item \textbf{Ground Truth:} Same colorized segmentation overlaid with the
        oracle verdict.  When a skeleton file is provided, skeleton nodes
        belonging to the axons passing through Fragments~A or~B are highlighted
        in bright yellow, and a dashed line marks the oracle crossing for
        positive pairs.  A large badge labels the outcome: \textbf{TP}
        (correct accept), \textbf{FP} (false accept), \textbf{FN} (missed
        merge), or \textbf{TN} (correct reject), together with an explicit
        ``Oracle: SHOULD MERGE'' or ``Oracle: NO MERGE'' annotation.
\end{enumerate}

This layout gives a human inspector all information needed to assess whether
the pipeline's decision is correct: the tissue context (Panel~1), the
pipeline's geometric reasoning (Panel~2), and the ground-truth verdict with
skeleton evidence (Panel~3).  Same-label candidates (two fragments from the
same segment, which trivially score 1.0) are excluded from sampling to avoid
contaminating the display with degenerate cases.
